The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2012-01-26 smurf_paper.tex
add histogram plot to cosmology section

The first step, map generation, is different from that described in
Section~\ref{sec:point} in two key ways. Since the locations of
sources are unknown \emph{a priori}, a map constraint is not
employed. Large-scale diverging structures in the map must be
mitigated, and the method used in this example (the default processing
in SMURF) is to apply a high-pass filter to the data \emph{once} as a
pre-processing steps. The iterative solution is then run using only
\model{COM}, \model{EXT}, \model{AST}, and \model{NOI}. In other
words, there is no information in the bolometer signals below some
cutoff frequency, and residual noise is removed only using iterative
estimation of the residual (higher-frequency) common-mode. The map is
shown in Fig.~\ref{fig:lockman_maps}a. The map-maker has been tested
in two ways: (i) large numbers of iterations are used to verify that
the maps converge without the growth of large structure; and (ii)
adding synthetic sources to the data at a range of brightnesses verify
that the map-maker response to them is \emph{linear} (i.e. the
relative shape and amplitude compared to the input source is
independent of brightness). The results of this second test are shown
by the solid and dotted lines in Fig.~\ref{fig:psf},
respectively. Clearly the down-side of using a high-pass filter as a
pre-processing step, and having no other map-constraints, has the
down-side of introducing sidelobes around the main peak. ... high-pass
filter + iterative common-mode removal takes place of PCA cleaning in
AzTEC reduction ... probably more linear.

Even though the map \emph{looks} extremely flat, there is a mixture of
faint astronomical sources, and what is probably residual
low-frequency noise causing faint ripples visible to the naked
eye. Since the mixture of the two components is unknown, the first
step is to remove the low-frequency noise in a way that does not
affect (appreciably) the real sources. To accomplish this goal, the
map is run through a whitening filter that ideally results in a flat
angular noise power spectrum. The noise power spectrum is estimated
from a \emph{jackknife map}: maps are produced from two independent
halves of the total data set, and then a pixel-by-pixel difference is
calculated. Specifically, the jackknife signal in a map pixel,
$S_\mathrm{JK}$, and variance, $\sigma^2_\mathrm{JK}$ are estimated
from the map quantities, $S_1$, $S_2$, $\sigma^2_1$, and $\sigma^2_2$
as,
%
\begin{eqnarray}
S_\mathrm{JK} = \frac{S_1 - S_2}{2} \\
\sigma^2_\mathrm{JK} = \frac{\sigma^2_1 + \sigma^2_2}{2}
\end{eqnarray}

Provided that the noise in one half of the data is uncorrelated with
that from the other half, the noise in the jackknife map should
resemble that of the raw map. The astronomical signal, however, should
be cleanly removed (provided that there are no strong time-varying
signals, and also assuming that errors due to calibration between the
two halves are insignificant). The approach we have taken to minimize
systematics is to produce two halves of the jackknife using odd and
even scan numbers (i.e., each map will contain an even sampling of
data across the full dataset).

@@

>>>>

The first step, map generation, is different from that described in
Section~\ref{sec:point} in two key ways. Since the locations of
sources are unknown \emph{a priori}, a map constraint is not
employed. Large-scale diverging structures in the map must be
mitigated, and the method used in this example (the default processing
in SMURF) is to apply a high-pass filter to the data \emph{once} as a
pre-processing steps. The iterative solution is then run using only
\model{COM}, \model{EXT}, \model{AST}, and \model{NOI}. In other
words, there is no information in the bolometer signals below some
cutoff frequency, and residual noise is removed only using iterative
estimation of the residual (higher-frequency) common-mode. The map is
shown in Fig.~\ref{fig:lockman_maps}a. The map-maker has been tested
in two ways: (i) large numbers of iterations are used to verify that
the maps converge without the growth of large structure; and (ii)
adding synthetic sources to the data at a range of brightnesses verify
that the map-maker response to them is \emph{linear} (i.e. the
relative shape and amplitude compared to the input source is
independent of brightness). The results of this second test are shown
by the solid and dotted lines in Fig.~\ref{fig:lockman_psf},
respectively. Clearly the down-side of using a high-pass filter as a
pre-processing step, and having no other map-constraints, has the
down-side of introducing sidelobes around the main peak. ... high-pass
filter + iterative common-mode removal takes place of PCA cleaning in
AzTEC reduction ... probably more linear.

Even though the map \emph{looks} extremely flat, there is a mixture of
faint astronomical sources, and what is probably residual
low-frequency noise causing faint ripples visible to the naked
eye. Since the mixture of the two components is unknown, the first
step is to remove the low-frequency noise in a way that does not
affect (appreciably) the real sources. To accomplish this goal, the
map is run through a whitening filter that ideally results in a flat
angular noise power spectrum. The noise power spectrum is estimated
from a \emph{jackknife map}: maps are produced from two independent
halves of the total data set, and then a pixel-by-pixel difference is
calculated. Specifically, the jackknife signal in a map pixel,
$S_\mathrm{JK}$, and variance, $\sigma^2_\mathrm{JK}$ are estimated
from the map quantities, $S_1$, $S_2$, $\sigma^2_1$, and $\sigma^2_2$
as,
%
\begin{eqnarray}
S_\mathrm{JK} = \frac{S_1 - S_2}{2} \\
\sigma^2_\mathrm{JK} = \frac{\sigma^2_1 + \sigma^2_2}{4}
\end{eqnarray}

Provided that the noise in one half of the data is uncorrelated with
that from the other half, the signal in the jackknife map should
resemble noise drawn from the same parent distribution as that of the
real map. The astronomical signal, however, should be cleanly removed
(provided that there are no strong time-varying signals, and also
assuming that errors due to calibration between the two halves are
insignificant). The approach we have taken to minimize systematics is
to produce two halves of the jackknife using odd and even scan numbers
(i.e., each map will contain a nearly evenly-spaced sampling of data
across the full dataset).

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{lockman_hist.pdf}
\caption{Histograms of S/N using pixels from the central region of
  Lockman Hole (second contour) in Fig.~\ref{fig:lockman_maps}a. The
  top panel shows the histograme for pixels from the whitened signal
  and jackknife maps (Figs.~\ref{fig:lockman_maps}b,e), compared to a
  Gaussian distribution with mean zero, standard deviation one, and an
  area normalized to the total number of pixels -- the expected
  distribution for a map of spatially-uncorrelated Gaussian noise. The
  good agreement indicates that these maps are indeed dominated by
  white noise, whose amplitude has been accurately estimated during
  map-making. The lower panel shows the results for matched-filtered
  signal and jackknife maps (Figs.~\ref{fig:lockman_maps}c,f). The
  filtered jackknife map distribution is still close to the
  expectation (Gaussian), but now the matched-filter has picked out
  significant signal leading to the large positive tail. }
\label{fig:lockman_hist}
\end{figure}
